\section{Architecture}

Given the project goals, the architectures structure, so the architectures indivudal components, units, are largely already given. 

  The requirement for turing completeness, also known as computationally universal, so the requirement to be able to exectue any algorithm \cite{cit.needed}, poses several restrictions. First of all, a computationally universal cannot exist, HALTING PROBLEMB

  SO generally defined as:
  
  \cite{cit.needed}


difference in bus lengths, 8-bit data cannot be infinite can it?
,
The Von-Neumann-Architecture \cite{vonneuman1945a} intends that the architecture is split up into 
alu
cu 
memory 
I/O
bus 

Given the requiremenmt for 8bit and infinite memory, a deviation is required. 

data and addr bus will be split to a certainn extent. in a way the first 8bit of the data bus can be latched onto the addr bus\dots
finally also the control bus will be external, as this is once again much wider. also from the component layout a control bus that is the same as the data bus would still make sense. 

this however does not limit the architecture to still follow the single bus and data word at a time. (which is why the VNA is still promiinent today. no race conditions)

Additionally there is no problem adhering to the VNC. 


Onto bus at rising clock
from bus at falling clock


\input{2-mainpart/architecture/alu}

\input{2-mainpart/architecture/memory}

\input{2-mainpart/architecture/register}

\input{2-mainpart/architecture/control_unit}

\input{2-mainpart/architecture/bus}

\input{2-mainpart/architecture/input_output}
